---
title: "Initial Korean covid 19"
author: "lana"
date: "todays date"
output:
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(digits=2)
require(tidyverse)
require(tigerstats)
require(lubridate)
kRdata <- read.csv(file = 'kRdata.csv')
kRdata <- kRdata %>%
mutate(across(where(is.character), as.factor))

```
### Data is from the Patient_Info dataset from South Korea 1 Jan to 1 May 2020
### These data were retrieved from a downloadable zip file from kaggle

Some of the data was uncoded for sex - sex must be just male or female
77 of the data entries were coded sex  as blank, so we omit these.
state of a patient must be died or alive - we adjust that as well   
  
```{r}
str(kRdata)
```
  
  
  
  
# Your Question - Is dying independent of sex?
$H_0$: there is no difference in dying based on sex

$H_A$: there is no difference in dying based on sex


#Methods

the variable sex is categorical and the varibale results are categorical, so therefore we use cat~cat

## Graphical results - use ggplot commands to do this  
```{r}
dd2 <- kRdata %>% group_by(result,Sex) %>% summarize(count=n())
mutate(prcnt=count/sum(count))


#the group_by followed by summarize(count=n())

basicC <-ggplot(dd2,aes(x=Sex,y=count,fill=result))
basicC + geom_bar(stat="identity", position="dodge") 

#Now for percentage plot 

basicCC <- ggplot(dd2,aes(x=Sex,y=prcnt*100,fill=result))
basicCC + geom_bar(stat="identity", position = "dodge")
```

```{r}
#Load the ggplot2 package
library(ggplot2)

# Create a group bar chart
ggplot(kRdata, aes(x = result, fill = Sex)) +
  geom_bar(position = "dodge") +
  labs(x = "Result", y = "Sex") +
  ggtitle("Grouped Bar Chart of Sex and Result")

```

this bar chart shows the percentage of male and female who died from covid, and shows people who had it but are still alive 
  
## Numerical results
```{r}
table1 <- xtabs(~result + Sex, data=kRdata)
rowPerc(table1)
colPerc(table1)
```
the table shows the percentage of male and female who had covid that is alive and the percentage of who died. male survivors percentage is 44% while female survivors is 56%. 

## Inferential results 

```{r}
chisq.test(table1)
chisqtestGC(table1)
fisher.test(table1)
```

the table shows the p value in the chi-square stat, which is 0.0012. the p value in the fisher test, is 0.0009. therefore we reject both null hypothesis. 

## Conclusions

the question is "is dying the independent of sex". in conclusion more females that stayed alive covid than males. male were more prone to getting covid then women. the p value was 0.0012 and 0.0009, which had us reject the null hypothesis. 

